### 2017-03-30 scores/score.tex

\definecolor{mygrey}{RGB}{100,100,100}
\draw[line width=1pt] (0, 17.1595)  -- (0, 15.6265) ; % TimeNode.0 
\draw[line width=1pt] (1.64502, 17.1595)  -- (1.64502, 17.1595) ; % drag29wine94 
\draw[line width=1pt] (5, 17.1595)  -- (5, 17.1595) ; % slid38maya79 
\draw[line width=1pt] (1.47186, 15.6265)  -- (1.47186, 15.6265) ; % glut76duff98 
\draw[line width=1pt] (4.85554, 15.6265)  -- (4.85554, 15.6265) ; % ouch30vade32 
\draw[dashed,line width=1pt] (0, 17.1595)  -- (1.64502, 17.1595) ; % gaur3skin98 
\draw[line width=1pt] (1.64502, 17.1595)  -- (5, 17.1595) ; % f1 
\draw[line width=1pt] (1.64502, 17.0595) rectangle (5, 16.0595) node[pos=.5] {$f_1$};
\draw[dashed,line width=1pt] (0, 15.6265)  -- (1.47186, 15.6265) ; % mail74chic13 
\draw[line width=1pt] (1.47186, 15.6265)  -- (4.85554, 15.6265) ; % f2 
\draw[line width=1pt] (1.47186, 15.5265) rectangle (4.85554, 14.5265) node[pos=.5] {$f_2$};

\fill[color=mygrey] (0, 17.1595) circle (0.075) ; % State.0 
\fill[color=mygrey] (1.64502, 17.1595) circle (0.075) ; % State.1 
\fill[color=mygrey] (5, 17.1595) circle (0.075) ; % State.2 
\fill[color=mygrey] (0, 15.6265) circle (0.075) ; % State.3 
\fill[color=mygrey] (1.47186, 15.6265) circle (0.075) ; % State.4 
\fill[color=mygrey] (4.85554, 15.6265) circle (0.075) ; % State.5